---
title: "ImageAlignR"
date: "`r format(Sys.time(), '%B %d, %Y')`"
bibliography: Refs.bib
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo=FALSE}
biblio <- bibtex::read.bib("Refs.bib")
```
[![CRAN status](https://www.r-pkg.org/badges/version/ImageAlignR)](https://cran.r-project.org/package=ImageAlignR)
[![lifecycle](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
[![Travis build status](https://travis-ci.org/srvanderplas/ImageAlignR.svg?branch=master)](https://travis-ci.org/srvanderplas/ImageAlignR)
[![Coverage status](https://codecov.io/gh/srvanderplas/ImageAlignR/branch/master/graph/badge.svg)](https://codecov.io/github/srvanderplas/ImageAlignR?branch=master)

# Image Alignment Methods:

1. Single-Image Alignment
    - Ellipse-based: Align an image so that its main axes are the y and x axis of the cartesian coordinate system. Rotates an image so that it is straight up-and-down.    
    `r capture.output(print(biblio["gwoShoeprintRetrievalCore2016"]))`
2. Dual-Image Alignment
    - Keypoint Alignment: Keypoint detection using the following steps:
        1. Harris corner detector
        2. Calculate key image orientations
        3. Feature detection
        4. Point matching using KNN to find similar points
        5. RANSAC to get a consistent homology
        6. Image transformation

